\documentclass[12pt]{article}
\usepackage{url,graphicx,tabularx,array}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%% PREAMBLE %%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% -------------------- defaults -------------------- %
% load lots o' packages

% layout control
\usepackage[paper=a4paper,left=25mm,right=25mm,top=20mm,bottom=25mm]{geometry}
\usepackage[onehalfspacing]{setspace}
\setlength{\parskip}{.5em}
\usepackage{rotating}
\usepackage{setspace}

% math typesetting
\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}

\usepackage{verbatim}
\usepackage{listings}

% restricts float objects to be inserted before end of section
% creates float barriers
\usepackage[section]{placeins}

% tables
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{longtable}

\usepackage[%
decimalsymbol=.,
digitsep=fullstop
]{siunitx}

% to adapt caption style
\usepackage[font={small},labelfont=bf]{caption}

% references
\usepackage[longnamesfirst]{natbib}
\bibpunct{(}{)}{;}{a}{}{,}

% footnotes at bottom
\usepackage[bottom]{footmisc}

% to change enumeration symbols begin{enumerate}[(a)]
\usepackage{enumerate}

% to make enumerations and itemizations within paragraphs or
% lines. f.i. begin{inparaenum} for (a) is (b) and (c)
\usepackage{paralist}

% to colorize links in document. See color specification below
\usepackage[pdftex,hyperref,x11names]{xcolor}

% for multiple references and insertion of the word "figure" or "table"
% \usepackage{cleveref}

% load the hyper-references package and set document info
\usepackage[pdftex]{hyperref}

% graphics stuff
\usepackage{coordsys}
\usepackage{tikz}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage[space]{grffile} % allows us to specify directories that have spaces
\usepackage{placeins} % prevents floats from moving past a \FloatBarrier
%\usepackage{tikz}
% \usepackage{pgfplots}

% define clickable links and their colors
\hypersetup{%
unicode=false,          % non-Latin characters in Acrobat's bookmarks
pdftoolbar=true,        % show Acrobat's toolbar?
pdfmenubar=true,        % show Acrobat's menu?
pdffitwindow=false,     % window fit to page when opened
pdfstartview={FitH},    % fits the width of the page to the window
pdfnewwindow=true,%
pagebackref=false,%
pdfauthor={Matt Dickenson},%
pdftitle={Title},%
colorlinks,%
citecolor=black,%
filecolor=black,%
linkcolor=black,%
urlcolor=RoyalBlue4}

\lstset{
    basicstyle=\ttfamily,
    upquote=true,
    breaklines=true,
    tabsize=2,
    postbreak=\raisebox{0ex}[0ex][0ex]{\ensuremath{\hookrightarrow}},
    breakatwhitespace=true,
    numbers=left,
    showstringspaces=false
}

% -------------------------------------------------- %



%-- Set options
\setlength{\parskip}{1ex} %--skip lines between paragraphs
\setlength{\parindent}{0pt} %--don't indent paragraphs
\newcommand{\inputTikZ}[2]{%  
     \scalebox{#1}{\input{#2}}  
}


%-- Commands for header
\renewcommand{\title}[1]{\textbf{#1}\\}
\renewcommand{\line}{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}\hline\\\end{tabularx}\\[-0.5cm]}
\newcommand{\leftright}[2]{\begin{tabularx}{\textwidth}{X>{\raggedleft}X}#1%
& #2\\\end{tabularx}\\[-0.5cm]}
%\linespread{2} %-- Uncomment for Double Space
\begin{document}

% ----------------------TITLE----------------------- %
\title{CS 201 - HW 3 WriteUp}
\line
\leftright{\today}{Matt Dickenson} %-- left and right positions in the header
\setlength{\parindent}{16pt}

\section*{Part A}

\paragraph{1)} The time to generate order-5 text using \texttt{romeo.txt} is shown in the table below:

\begin{table}[h]
\begin{center}
\begin{tabular}{lr}
Characters & Time (seconds) \\
\hline
100 & 0.042 \\
200 & 0.083 \\
400 & 0.162 \\
800 & 0.338 \\
1600 & 0.641 \\
\end{tabular}
\end{center}
\end{table}

As the table shows, doubling the size of the requested output approximately doubles the size of the time it takes to generate the text, plus a fixed amount of time of about 2 milliseconds. Thus, the brute force model is $O(n)$where $n$ is the size of the output.  

An order-10 model with this same textual input takes about the same amount of time, but an order-1 model seems to take about 1.4 to 1.5 times as long at all sizes of input. This is likely because the number of 1-grams is much larger than the number of 5- or 10-grams in the text. 

\paragraph{2)} The size of the input affects the run time through the \texttt{while} loop on line 71 of \texttt{MarkovModel.java}. Since the run time of nested loops are multiplicative, the run time is a function of $n \times m$ where $n$ is the length of the requested output and $m$ is the length of the input text. In (A1) we held $m$ fixed. If we change the input text from \emph{Romeo and Juliet} to \emph{Scarlet Letter} (which is over 3 times as long), I would expect the runtime time to increase by $3n$. 

Consequently, I predict that generating order-5 text of lengths 400, 800, and 1600 characters would take about 0.5, 1.0, and 1.8 seconds, respectively. If we were to use the King James Bible text, $m$ would be roughly 9 times as long as it is with \emph{Scarlet Letter} and about 30 times as long as \emph{Romeo and Juliet}. I predict that generating 1600 characters of text with an order-5 model using the King James corpus would take about 16.2 ($9\times1.8$) seconds. 

\paragraph{3)} Using the Map method in \texttt{MapMarkovModel.java}, the run times are as follows for an order-5 model with \texttt{romeo.txt}:

\begin{table}[h]
\begin{center}
\begin{tabular}{lr}
Characters & Time (seconds) \\
\hline
100 & 0.001 \\
200 & 0.001\\
400 & 0.003 \\
800 & 0.008 \\
1600 & 0.014 \\
\end{tabular}
\end{center}
\end{table}

This data indicates to me that the Map method is still $O(n)$ because the run time approximately doubles as $n$ doubles. However, it does reduce the time required by a factor of about 40. This is one weakness of Big-Oh notation: sometimes constant factors make a big difference! 

\end{document}
